---
title: "snRNA_spleen_part1"
output: html_document
date: "2024-07-04"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### I - Setting the working environment
#Loading libraries for the analysis
```{r}
library(Seurat)
library(tidyr)
library(purrr)
library(dplyr)
library(reshape2)
library(ggplot2)
library(qs)
library(SoupX)
library(cowplot)
library(harmony)
library(scCustomize)
```

#Defining the working folder 
```{r}
setwd("~/Thèse/SMP_ST/Data/snRNAseq/Seurat")
```

### II - Filtering and Quality Control
#Using SoupX to get rid of ambient RNA
#Small adaptation of the original function to deal with CellRanger multi-out FLEX assays specificities (different number of genes between raw and filtered matrices)
```{r}
Create_soup_and_correct = function(name_folder){
  #defining the folder containing the CellRanger outputs 
  root= "~snRNAseq/per_sample_outs"
  filtered_toc <- Read10X(data.dir = file.path(root, name_folder,"count","sample_filtered_feature_bc_matrix"))
  raw_tod <- Read10X(data.dir = file.path(root, name_folder,"count","sample_raw_feature_bc_matrix"))
  
  filt_genes <- rownames(filtered_toc)
  # Subset raw.matrix to keep only the genes in filt.matrix
  raw_tod_subset <- raw_tod[rownames(raw_tod) %in% filt_genes, ]
  #Get the cluster annotation
  mDat = NULL
  tgt =  file.path(root,name_folder,'count','analysis','clustering', 'gene_expression_graphclust','clusters.csv')
               
  if(file.exists(tgt)){
    clusters = read.csv(tgt)
    mDat = data.frame(clusters=clusters$Cluster,row.names=clusters$Barcode)}
  #Add fine grained clusters too if present
  tgt = file.path(root,name_folder,'count','analysis','clustering','gene_expression_kmeans_10_clusters','clusters.csv')
  
  if(file.exists(tgt)){
    clusters = read.csv(tgt)
    mDat$clustersFine = clusters$Cluster }
  #Get tSNE if available and point to it
  print("loading tsne information")
  tgt = file.path(root,name_folder,'count','analysis','tsne','gene_expression_2_components','projection.csv')
               
  if(file.exists(tgt)){
    tsne = read.csv(tgt)
    print("tgt csv object loaded")
    mDat$tSNE1 = tsne$TSNE.1[match(rownames(mDat),tsne$Barcode)]
    mDat$tSNE2 = tsne$TSNE.2[match(rownames(mDat),tsne$Barcode)]
    DR = c('tSNE1','tSNE2')
    print("DR created") }

  #Ensure rownames of metadata match column names of counts
  print(typeof(filtered_toc))
  print(colnames(filtered_toc))
  if(any(rownames(mDat)!=colnames(filtered_toc))) { 
    print("Error matching meta-data to cell names.")}
  
  print("creating soup object")
  sc = SoupChannel(raw_tod_subset, filtered_toc, metaData = mDat, DR=DR, calcSoupProfile = FALSE)
  sc = estimateSoup(sc)

  print("correcting and returning")
  # Estimate rho
  sc = autoEstCont(sc)
  # Clean the data
  out_to_return = adjustCounts(sc)
  
  return(out_to_return)}
```

#Creating Seurat objects using the corrected matrices
```{r}
out_S06_soup = Create_soup_and_correct("06H11604")
S06H11604 = CreateSeuratObject(out_S06_soup, project = "S06")

out_T044_soup = Create_soup_and_correct("0440181")
T044 = CreateSeuratObject(out_T044_soup, project = "T044")

out_H0409053_soup = Create_soup_and_correct("H0409053")
H0409053 = CreateSeuratObject(out_H0409053_soup, project = "H040")
  
out_S1306559RA_soup = Create_soup_and_correct("S1306559RA")
S1306559RA = CreateSeuratObject(out_S1306559RA_soup, project = "S13")
  
out_S1602156Z03_soup = Create_soup_and_correct("S1602156Z03")
S1602156Z03 = CreateSeuratObject(out_S1602156Z03_soup, project = "S16_156")
  
out_S1726736R14_soup = Create_soup_and_correct("S1726736R14")
S1726736R14 = CreateSeuratObject(out_S1726736R14_soup, project = "S17")

out_S1902552Z04_soup = Create_soup_and_correct("S1902552Z04")
S1902552Z04 = CreateSeuratObject(out_S1902552Z04_soup, project = "S19")

out_S2213861_soup = Create_soup_and_correct("S2213861-1")
S2213861 = CreateSeuratObject(out_S2213861_soup, project = "S22")

list_sn_corrected_seurat_object = list(S06H = S06H11604, T044 = T044, H040 = H0409053,  S13 = S1306559RA,  S16_156 = S1602156Z03, S17 = S1726736R14, S19 = S1902552Z04, S22 = S2213861)
```

#Creating the merged Seurat object containing information from all samples
```{r}
for (object in names(list_sn_corrected_seurat_object)) {
  list_sn_corrected_seurat_object[[object]][["percent.mito"]] = PercentageFeatureSet(list_sn_corrected_seurat_object[[object]], pattern = "^MT-")
  list_sn_corrected_seurat_object[[object]] = NormalizeData(list_sn_corrected_seurat_object[[object]])}

corrected_merged_object_sn_8patients = scCustomize::Merge_Seurat_List(list_sn_corrected_seurat_object, add.cell.ids = c("S06H","T044","TH040","S13","S16_156","S17","S19","S22"))
```

#Removing doublets with scDblFinder
https://bioconductor.org/packages/release/bioc/vignettes/scDblFinder/inst/doc/scDblFinder.html
```{r}
library(scDblFinder)
library(BiocParallel)

scDblFinder_filtered_corrected_merged_object_sn_8patients <- subset(corrected_merged_object_sn_8patients, subset = nCount_RNA > 150 )

#this packages works with objects in the SingleCellExperiment format
soupx_corrected.merged.sce <- as.SingleCellExperiment(scDblFinder_filtered_corrected_merged_object_sn_8patients)
sce <- scDblFinder(soupx_corrected.merged.sce, samples="orig.ident", dbr.sd= 1)
table(sce$scDblFinder.class)

#transfer results back into the Seurat Object
scDblFinder_filtered_corrected_merged_object_sn_8patients$scDblFinder.class <- sce$scDblFinder.class
scDblFinder_filtered_corrected_merged_object_sn_8patients$scDblFinder.weighted <- sce$scDblFinder.weighted
scDblFinder_filtered_corrected_merged_object_sn_8patients$scDblFinder.score <- sce$scDblFinder.score
scDblFinder_filtered_corrected_merged_object_sn_8patients$scDblFinder.cxds_score <- sce$scDblFinder.cxds_score
table(scDblFinder_filtered_corrected_merged_object_sn_8patients$scDblFinder.class)
```

#QC filtering - Post SoupX and post FindDblt
```{r}
plot_grid(ncol = 3,RidgePlot(scDblFinder_filtered_corrected_merged_object_sn_8patients,features = "percent.mito", group.by = "orig.ident")  + theme(axis.title.y.right = element_text(size = 15), legend.position = 'none'),
RidgePlot(scDblFinder_filtered_corrected_merged_object_sn_8patients,features = "nCount_RNA", group.by = "orig.ident")  + theme(axis.title.y.right = element_text(size = 15), legend.position = 'none') + xlim(c(0,4000)),
RidgePlot(scDblFinder_filtered_corrected_merged_object_sn_8patients,features = "nFeature_RNA", group.by = "orig.ident")   + theme(axis.title.y.right = element_text(size = 15), legend.position = 'none')  + xlim(c(0,4000)) )

#identifying potential suspect clusters before removing doublets using scDbltFinder class information
scDblFinder_filtered_corrected_merged_object_sn_8patients = FindClusters(scDblFinder_filtered_corrected_merged_object_sn_8patients, res = 1)
scDblFinder_filtered_corrected_merged_object_sn_8patients <- RunUMAP(scDblFinder_filtered_corrected_merged_object_sn_8patients)
DimPlot(scDblFinder_filtered_corrected_merged_object_sn_8patients)
FeaturePlot(scDblFinder_filtered_corrected_merged_object_sn_8patients, features = "scDblFinder.score") +  scale_color_viridis_c()

scDblFinder_filtered_corrected_merged_object_sn_8patients@meta.data[,c("RNA_snn_res.1","scDblFinder.score")] %>% group_by(RNA_snn_res.1) %>%   summarize(n_score = n(),
              n_dblts = sum(scDblFinder.score > 0.5),
              pct_dblts = n_dblts / n_score)

#QC filtering
final_filtered_corrected_merged_object_sn_8patients <- subset(scDblFinder_filtered_corrected_merged_object_sn_8patients, subset = nFeature_RNA > 150 & nFeature_RNA < 1500 & percent.mito < 5 & scDblFinder.class == "singlet"  & RNA_snn_res.1 != 3)

plot_grid(ncol = 3,RidgePlot(final_filtered_corrected_merged_object_sn_8patients,features = "percent.mito", group.by = "orig.ident")  + theme(axis.title.y.right = element_text(size = 15), legend.position = 'none'),
RidgePlot(final_filtered_corrected_merged_object_sn_8patients,features = "nCount_RNA", group.by = "orig.ident")  + theme(axis.title.y.right = element_text(size = 15), legend.position = 'none') + xlim(c(0,4000)),
RidgePlot(final_filtered_corrected_merged_object_sn_8patients,features = "nFeature_RNA", group.by = "orig.ident")   + theme(axis.title.y.right = element_text(size = 15), legend.position = 'none')  + xlim(c(0,2000)) )
```

### III - Integration and cells annotation
#Looking at data without integration
```{r}
final_filtered_corrected_merged_object_sn_8patients <- FindVariableFeatures(final_filtered_corrected_merged_object_sn_8patients, selection.method = "vst", nfeatures = 2000)

final_filtered_corrected_merged_object_sn_8patients <- ScaleData(final_filtered_corrected_merged_object_sn_8patients, features = rownames(final_filtered_corrected_merged_object_sn_8patients))

final_filtered_corrected_merged_object_sn_8patients <- RunPCA(final_filtered_corrected_merged_object_sn_8patients, features = VariableFeatures(object = final_filtered_corrected_merged_object_sn_8patients))
ElbowPlot(final_filtered_corrected_merged_object_sn_8patients)

final_filtered_corrected_merged_object_sn_8patients <- FindNeighbors(final_filtered_corrected_merged_object_sn_8patients, dims = 1:10)
final_filtered_corrected_merged_object_sn_8patients <- FindClusters(final_filtered_corrected_merged_object_sn_8patients, resolution = 0.5)

final_filtered_corrected_merged_object_sn_8patients <- RunUMAP(final_filtered_corrected_merged_object_sn_8patients, dims = 1:10)
DimPlot(final_filtered_corrected_merged_object_sn_8patients, reduction = "umap", group.by = "orig.ident")
```

#Integrating samples using Harmony
```{r}
ndim = 20
set.seed(2)

filtered_corrected_merged_object_sn_8patients_harmony <- RunHarmony(final_filtered_corrected_merged_object_sn_8patients, group.by.vars = "orig.ident")
filtered_corrected_merged_object_sn_8patients_harmony <- RunUMAP(filtered_corrected_merged_object_sn_8patients_harmony, reduction = "harmony", dims = 1:ndim, seed.use = 2)
DimPlot(filtered_corrected_merged_object_sn_8patients_harmony, reduction = "umap")
DimPlot(filtered_corrected_merged_object_sn_8patients_harmony, reduction = "umap", group.by = "orig.ident")

filtered_corrected_merged_object_sn_8patients_harmony <- RunTSNE(filtered_corrected_merged_object_sn_8patients_harmony, reduction = "harmony", dims = 1:ndim)
filtered_corrected_merged_object_sn_8patients_harmony <- FindNeighbors(filtered_corrected_merged_object_sn_8patients_harmony, reduction = "harmony", dims = 1:ndim) %>% FindClusters(resolution = 1.5)
DimPlot(filtered_corrected_merged_object_sn_8patients_harmony, reduction = "umap")
```

```{r}
sample_levels <- c("H040","T044","S17","S22","S16_156","S13","S19","S06") 
filtered_corrected_merged_object_sn_8patients_harmony$orig.ident = factor(filtered_corrected_merged_object_sn_8patients_harmony$orig.ident, levels= sample_levels)

#Green colors for healthy samples
green_palette <- RColorBrewer::brewer.pal(3, "Greens")[2:3]
#Oranges for Disease
orange_palette <- RColorBrewer::brewer.pal(6, "Oranges")[1:6]
#Final colors
sample_colors <- setNames(c(green_palette, orange_palette), sample_levels)
sample_colors[3] <- "#DAA520"  # Goldenrod (ocre/or)

#QC plots with specific colors
p1 <- VlnPlot(filtered_corrected_merged_object_sn_8patients_harmony, features = "percent.mito", group.by = "orig.ident", pt.size = 0) + xlab("") + scale_fill_manual(values = sample_colors) +  theme(axis.title.y.right = element_text(size = 15), legend.position = 'none') + ylim(c(0, 10)) + geom_boxplot()
p2 <- VlnPlot(filtered_corrected_merged_object_sn_8patients_harmony, features = "nCount_RNA", group.by = "orig.ident", pt.size = 0) + xlab("") +scale_fill_manual(values = sample_colors) + theme(axis.title.y.right = element_text(size = 15), legend.position = 'none') + ylim(c(0, 3000)) +geom_boxplot()
p3 <- VlnPlot(filtered_corrected_merged_object_sn_8patients_harmony,     features = "nFeature_RNA", group.by = "orig.ident", pt.size = 0) + xlab("") + scale_fill_manual(values = sample_colors) + theme(axis.title.y.right = element_text(size = 15), legend.position = 'none') +  ylim(c(0, 2000)) + geom_boxplot()

p1+p2+p3

harmony_plot = DimPlot(filtered_corrected_merged_object_sn_8patients_harmony, group.by = "orig.ident", split.by = "orig.ident") +  scale_color_manual(values = sample_colors)
ggsave("~figures/harmony_plot.png", harmony_plot)
```

#Cluster annotation
```{r}
filtered_corrected_merged_object_sn_8patients_harmony@active.ident = filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5
markers = FindAllMarkers(filtered_corrected_merged_object_sn_8patients_harmony, only.pos = TRUE)
Dimplot(filtered_corrected_merged_object_sn_8patients_harmony, group.by = "RNA_snn_res.1.5")

filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation= "Annotation"

filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 0] = "Lymphocytes 1"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 1] = "Lymphocytes 1"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 2] = "Sinusoidal Cells"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 3] = "Lymphocytes 2"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 4] = "Sinusoidal Cells"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 5] = "Lymphocytes 2"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 6] = "Macrophages"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 7] = "Fibro"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 8] = "Erythro"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 9] = "Monocytes"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 10] = "HSPC"

filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 11] = "Plasma Cells"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 12] = "Megakaryocytes"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 13] = "Neutrophils 1"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 14] = "Neutrophils 2"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 15] = "Megakaryocytes"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 16] = "Plasma Cells"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 17] = "Sinusoidal Cells"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 18] = "Erythro"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 19] = "Dendritic Cells (cDc1)"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 20] = "Fibro"

filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 21] = "Erythro"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 22] = "CE"

DimPlot(filtered_corrected_merged_object_sn_8patients_harmony, group.by = "RNA_snn_res.1.5")

filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation = factor(filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation, levels = c("CE","Sinusoidal Cells","Fibro","HSPC","Erythro","Megakaryocytes","Neutrophils 1","Neutrophils 2","Monocytes","Macrophages","Dendritic Cells (cDc1)","Lymphocytes 1","Lymphocytes 2","Plasma Cells"))

DimPlot(filtered_corrected_merged_object_sn_8patients_harmony, group.by = "simplified_annotation")
```

