---
title: "snRNA_spleen_part1"
output: html_document
date: "2024-07-04"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script includes all the steps required to process the Cell Ranger output matrices from the eight snRNA-seq samples and generate a final, clean, annotated Seurat object.

### I - Setting up the working environment
#Loading libraries for the analysis
```{r}
library(Seurat)
library(tidyr)
library(purrr)
library(dplyr)
library(reshape2)
library(ggplot2)
library(qs)
library(SoupX)
library(cowplot)
library(harmony)
library(scCustomize)
```

#Defining the working folder 
```{r}
setwd("~/Thèse/SMP_ST/Data/snRNAseq/Seurat")
```

### II - Filtering and Quality Control
#Using SoupX to get rid of ambient RNA
#Small adaptation of the original function to deal with CellRanger multi-out FLEX assays specificities (different number of genes between raw and filtered matrices)
```{r}
Create_soup_and_correct = function(name_folder){
  #defining the folder containing the CellRanger outputs 
  root= "~snRNAseq/per_sample_outs"
  filtered_toc <- Read10X(data.dir = file.path(root, name_folder,"count","sample_filtered_feature_bc_matrix"))
  raw_tod <- Read10X(data.dir = file.path(root, name_folder,"count","sample_raw_feature_bc_matrix"))
  
  filt_genes <- rownames(filtered_toc)
  # Subset raw.matrix to keep only the genes in filt.matrix
  raw_tod_subset <- raw_tod[rownames(raw_tod) %in% filt_genes, ]
  #Get the cluster annotation
  mDat = NULL
  tgt =  file.path(root,name_folder,'count','analysis','clustering', 'gene_expression_graphclust','clusters.csv')
               
  if(file.exists(tgt)){
    clusters = read.csv(tgt)
    mDat = data.frame(clusters=clusters$Cluster,row.names=clusters$Barcode)}
  #Add fine grained clusters too if present
  tgt = file.path(root,name_folder,'count','analysis','clustering','gene_expression_kmeans_10_clusters','clusters.csv')
  
  if(file.exists(tgt)){
    clusters = read.csv(tgt)
    mDat$clustersFine = clusters$Cluster }
  #Get tSNE if available and point to it
  print("loading tsne information")
  tgt = file.path(root,name_folder,'count','analysis','tsne','gene_expression_2_components','projection.csv')
               
  if(file.exists(tgt)){
    tsne = read.csv(tgt)
    print("tgt csv object loaded")
    mDat$tSNE1 = tsne$TSNE.1[match(rownames(mDat),tsne$Barcode)]
    mDat$tSNE2 = tsne$TSNE.2[match(rownames(mDat),tsne$Barcode)]
    DR = c('tSNE1','tSNE2')
    print("DR created") }

  #Ensure rownames of metadata match column names of counts
  print(typeof(filtered_toc))
  print(colnames(filtered_toc))
  if(any(rownames(mDat)!=colnames(filtered_toc))) { 
    print("Error matching meta-data to cell names.")}
  
  print("creating soup object")
  sc = SoupChannel(raw_tod_subset, filtered_toc, metaData = mDat, DR=DR, calcSoupProfile = FALSE)
  sc = estimateSoup(sc)

  print("correcting and returning")
  # Estimate rho
  sc = autoEstCont(sc)
  # Clean the data
  out_to_return = adjustCounts(sc)
  
  return(out_to_return)}
```

#Creating Seurat objects using the corrected matrices
```{r}
out_H1_soup = Create_soup_and_correct("H1")
H1 = CreateSeuratObject(out_H1_soup, project = "H1")

out_H2_soup = Create_soup_and_correct("H2")
H2 = CreateSeuratObject(out_H2_soup, project = "H2")

out_MPN1_soup = Create_soup_and_correct("MPN1")
MPN1 = CreateSeuratObject(out_MPN1_soup, project = "MPN1")

out_MPN2_soup = Create_soup_and_correct("MPN2")
MPN2 = CreateSeuratObject(out_MPN2_soup, project = "MPN2")

out_MPN6_soup = Create_soup_and_correct("MPN6")
MPN6 = CreateSeuratObject(out_MPN6_soup, project = "MPN6")
  
out_MPN7_soup = Create_soup_and_correct("MPN7")
MPN7 = CreateSeuratObject(out_MPN7_soup, project = "MPN7")

out_MPN8_soup = Create_soup_and_correct("MPN8")
MPN8 = CreateSeuratObject(out_MPN8_soup, project = "MPN8")

out_MPN9_soup = Create_soup_and_correct("MPN9")
MPN9 = CreateSeuratObject(out_MPN9_soup, project = "MPN9")

list_sn_corrected_seurat_object = list(H1 = H1,  H2 = H2,  MPN1 = MPN1, MPN2 = MPN2,  MPN6 = MPN6, MPN7 = MPN7, MPN8 = MPN8,  MPN9 = MPN9)
```

#Creating the merged Seurat object containing information from all samples
```{r}
for (object in names(list_sn_corrected_seurat_object)) {
  list_sn_corrected_seurat_object[[object]][["percent.mito"]] = PercentageFeatureSet(list_sn_corrected_seurat_object[[object]], pattern = "^MT-")
  list_sn_corrected_seurat_object[[object]] = NormalizeData(list_sn_corrected_seurat_object[[object]])}

corrected_merged_object_sn_8patients = scCustomize::Merge_Seurat_List(list_sn_corrected_seurat_object, add.cell.ids = c("H1","H2","MPN1","MPN2","MPN6","MPN7","MPN8","MPN9"))
```

#Removing doublets with scDblFinder
https://bioconductor.org/packages/release/bioc/vignettes/scDblFinder/inst/doc/scDblFinder.html
```{r}
library(scDblFinder)
library(BiocParallel)

scDblFinder_filtered_corrected_merged_object_sn_8patients <- subset(corrected_merged_object_sn_8patients, subset = nCount_RNA > 150 )

#this packages works with objects in the SingleCellExperiment format
soupx_corrected.merged.sce <- as.SingleCellExperiment(scDblFinder_filtered_corrected_merged_object_sn_8patients)
sce <- scDblFinder(soupx_corrected.merged.sce, samples="orig.ident", dbr.sd= 1)
table(sce$scDblFinder.class)

#transfer results back into the Seurat Object
scDblFinder_filtered_corrected_merged_object_sn_8patients$scDblFinder.class <- sce$scDblFinder.class
scDblFinder_filtered_corrected_merged_object_sn_8patients$scDblFinder.weighted <- sce$scDblFinder.weighted
scDblFinder_filtered_corrected_merged_object_sn_8patients$scDblFinder.score <- sce$scDblFinder.score
scDblFinder_filtered_corrected_merged_object_sn_8patients$scDblFinder.cxds_score <- sce$scDblFinder.cxds_score
table(scDblFinder_filtered_corrected_merged_object_sn_8patients$scDblFinder.class)
```

#QC filtering - Post SoupX and post FindDblt
```{r}
plot_grid(ncol = 3,RidgePlot(scDblFinder_filtered_corrected_merged_object_sn_8patients,features = "percent.mito", group.by = "orig.ident")  + theme(axis.title.y.right = element_text(size = 15), legend.position = 'none'),
RidgePlot(scDblFinder_filtered_corrected_merged_object_sn_8patients,features = "nCount_RNA", group.by = "orig.ident")  + theme(axis.title.y.right = element_text(size = 15), legend.position = 'none') + xlim(c(0,4000)),
RidgePlot(scDblFinder_filtered_corrected_merged_object_sn_8patients,features = "nFeature_RNA", group.by = "orig.ident")   + theme(axis.title.y.right = element_text(size = 15), legend.position = 'none')  + xlim(c(0,4000)) )

#identifying potential suspect clusters before removing doublets using scDbltFinder class information
scDblFinder_filtered_corrected_merged_object_sn_8patients <- FindVariableFeatures(scDblFinder_filtered_corrected_merged_object_sn_8patients, selection.method = "vst", nfeatures = 2000)
scDblFinder_filtered_corrected_merged_object_sn_8patients <- ScaleData(scDblFinder_filtered_corrected_merged_object_sn_8patients, features = rownames(scDblFinder_filtered_corrected_merged_object_sn_8patients))
scDblFinder_filtered_corrected_merged_object_sn_8patients <- RunPCA(scDblFinder_filtered_corrected_merged_object_sn_8patients, features = VariableFeatures(object = scDblFinder_filtered_corrected_merged_object_sn_8patients))
scDblFinder_filtered_corrected_merged_object_sn_8patients <- FindNeighbors(scDblFinder_filtered_corrected_merged_object_sn_8patients)
scDblFinder_filtered_corrected_merged_object_sn_8patients = FindClusters(scDblFinder_filtered_corrected_merged_object_sn_8patients, res = 1)
scDblFinder_filtered_corrected_merged_object_sn_8patients <- RunUMAP(scDblFinder_filtered_corrected_merged_object_sn_8patients)
DimPlot(scDblFinder_filtered_corrected_merged_object_sn_8patients)
FeaturePlot(scDblFinder_filtered_corrected_merged_object_sn_8patients, features = "scDblFinder.score") +  scale_color_viridis_c()

scDblFinder_filtered_corrected_merged_object_sn_8patients@meta.data[,c("RNA_snn_res.1","scDblFinder.score")] %>% group_by(RNA_snn_res.1) %>%   summarize(n_score = n(), n_dblts = sum(scDblFinder.score > 0.5), pct_dblts = n_dblts / n_score)

#QC filtering
final_filtered_corrected_merged_object_sn_8patients <- subset(scDblFinder_filtered_corrected_merged_object_sn_8patients, subset = nFeature_RNA > 150 & nFeature_RNA < 1500 & percent.mito < 5 & scDblFinder.class == "singlet"  & RNA_snn_res.1 != 3)

plot_grid(ncol = 3,RidgePlot(final_filtered_corrected_merged_object_sn_8patients,features = "percent.mito", group.by = "orig.ident")  + theme(axis.title.y.right = element_text(size = 15), legend.position = 'none'),
RidgePlot(final_filtered_corrected_merged_object_sn_8patients,features = "nCount_RNA", group.by = "orig.ident")  + theme(axis.title.y.right = element_text(size = 15), legend.position = 'none') + xlim(c(0,4000)),
RidgePlot(final_filtered_corrected_merged_object_sn_8patients,features = "nFeature_RNA", group.by = "orig.ident")   + theme(axis.title.y.right = element_text(size = 15), legend.position = 'none')  + xlim(c(0,2000)) )
```

### III - Integration and cells annotation
#Looking at data without integration
```{r}
final_filtered_corrected_merged_object_sn_8patients <- FindVariableFeatures(final_filtered_corrected_merged_object_sn_8patients, selection.method = "vst", nfeatures = 2000)

final_filtered_corrected_merged_object_sn_8patients <- ScaleData(final_filtered_corrected_merged_object_sn_8patients, features = rownames(final_filtered_corrected_merged_object_sn_8patients))

final_filtered_corrected_merged_object_sn_8patients <- RunPCA(final_filtered_corrected_merged_object_sn_8patients, features = VariableFeatures(object = final_filtered_corrected_merged_object_sn_8patients))
ElbowPlot(final_filtered_corrected_merged_object_sn_8patients)

final_filtered_corrected_merged_object_sn_8patients <- FindNeighbors(final_filtered_corrected_merged_object_sn_8patients, dims = 1:10)
final_filtered_corrected_merged_object_sn_8patients <- FindClusters(final_filtered_corrected_merged_object_sn_8patients, resolution = 0.5)

final_filtered_corrected_merged_object_sn_8patients <- RunUMAP(final_filtered_corrected_merged_object_sn_8patients, dims = 1:10)
DimPlot(final_filtered_corrected_merged_object_sn_8patients, reduction = "umap", group.by = "orig.ident")
```

#Integrating samples using Harmony to remove batch effect
```{r}
ndim = 20
set.seed(2)

filtered_corrected_merged_object_sn_8patients_harmony <- RunHarmony(final_filtered_corrected_merged_object_sn_8patients, group.by.vars = "orig.ident")
filtered_corrected_merged_object_sn_8patients_harmony <- RunUMAP(filtered_corrected_merged_object_sn_8patients_harmony, reduction = "harmony", dims = 1:ndim, seed.use = 2)
DimPlot(filtered_corrected_merged_object_sn_8patients_harmony, reduction = "umap")
DimPlot(filtered_corrected_merged_object_sn_8patients_harmony, reduction = "umap", group.by = "orig.ident")

filtered_corrected_merged_object_sn_8patients_harmony <- RunTSNE(filtered_corrected_merged_object_sn_8patients_harmony, reduction = "harmony", dims = 1:ndim)
filtered_corrected_merged_object_sn_8patients_harmony <- FindNeighbors(filtered_corrected_merged_object_sn_8patients_harmony, reduction = "harmony", dims = 1:ndim) %>% FindClusters(resolution = 1.5)
DimPlot(filtered_corrected_merged_object_sn_8patients_harmony, reduction = "umap")
```

```{r}
sample_levels <- c("H1","H2","MPN1","MPN2","MPN6","MPN7","MPN8","MPN9") 
filtered_corrected_merged_object_sn_8patients_harmony$orig.ident = factor(filtered_corrected_merged_object_sn_8patients_harmony$orig.ident, levels= sample_levels)

#Green colors for healthy samples
green_palette <- RColorBrewer::brewer.pal(3, "Greens")[2:3]
#Oranges for Disease
orange_palette <- RColorBrewer::brewer.pal(6, "Oranges")[1:6]
#Final colors
sample_colors <- setNames(c(green_palette, orange_palette), sample_levels)
sample_colors[3] <- "#DAA520"  # Goldenrod (ocre/or)

#QC plots with specific colors
p1 <- VlnPlot(filtered_corrected_merged_object_sn_8patients_harmony, features = "percent.mito", group.by = "orig.ident", pt.size = 0) + xlab("") + scale_fill_manual(values = sample_colors) +  theme(axis.title.y.right = element_text(size = 15), legend.position = 'none') + ylim(c(0, 10)) + geom_boxplot()
p2 <- VlnPlot(filtered_corrected_merged_object_sn_8patients_harmony, features = "nCount_RNA", group.by = "orig.ident", pt.size = 0) + xlab("") +scale_fill_manual(values = sample_colors) + theme(axis.title.y.right = element_text(size = 15), legend.position = 'none') + ylim(c(0, 3000)) +geom_boxplot()
p3 <- VlnPlot(filtered_corrected_merged_object_sn_8patients_harmony,     features = "nFeature_RNA", group.by = "orig.ident", pt.size = 0) + xlab("") + scale_fill_manual(values = sample_colors) + theme(axis.title.y.right = element_text(size = 15), legend.position = 'none') +  ylim(c(0, 2000)) + geom_boxplot()

p1+p2+p3

harmony_plot = DimPlot(filtered_corrected_merged_object_sn_8patients_harmony, group.by = "orig.ident", split.by = "orig.ident") +  scale_color_manual(values = sample_colors)
ggsave("~figures/harmony_plot.png", harmony_plot)
```

#Adding Metadata information on samples
```{r}
filtered_corrected_merged_object_sn_8patients_harmony$Disease <- "MPN" # or any other initialization value
filtered_corrected_merged_object_sn_8patients_harmony$Disease[filtered_corrected_merged_object_sn_8patients_harmony$orig.ident %in% c("H1","H2")] <- "Healthy"
```

#Cluster annotation
```{r}
filtered_corrected_merged_object_sn_8patients_harmony@active.ident = filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5
markers = FindAllMarkers(filtered_corrected_merged_object_sn_8patients_harmony, only.pos = TRUE)
Dimplot(filtered_corrected_merged_object_sn_8patients_harmony, group.by = "RNA_snn_res.1.5")

filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation= "Annotation"

filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 0] = "B Cells"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 1] = "B Cells"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 2] = "Sinusoidal Endothelial"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 3] = "T Cells"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 4] = "Sinusoidal Endothelial"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 5] = "T Cells"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 6] = "Macrophages"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 7] = "Reticular Cells"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 8] = "Ery"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 9] = "Monocytes"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 10] = "HSPC"

filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 11] = "Plasma Cells"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 12] = "Mk"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 13] = "Neutrophils 1"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 14] = "Neutrophils 2"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 15] = "Mk"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 16] = "Plasma Cells"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 17] = "Sinusoidal Endothelial"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 18] = "Ery"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 19] = "Dendritic Cells (cDc1)"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 20] = "Reticular Cells"

filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 21] = "Ery"
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 22] = "Classic Endothelial"

DimPlot(filtered_corrected_merged_object_sn_8patients_harmony, group.by = "RNA_snn_res.1.5")

filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation = factor(filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation, levels = c("Classic Endothelial","Sinusoidal Endothelial","Reticular Cells","HSPC","Ery","Mk","Neutrophils 1","Neutrophils 2","Monocytes","Macrophages","Dendritic Cells (cDc1)","B Cells","T Cells","Plasma Cells"))

DimPlot(filtered_corrected_merged_object_sn_8patients_harmony, group.by = "simplified_annotation")
```

#Detailed annotation (distinguishing subpopulation of reticular and hematopoietic cells)
```{r}
filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation= filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation

filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation = factor(filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation, levels = c("Classic Endothelial","Sinusoidal Endothelial","Reticular Cells","RP Reticular Cells","WP Reticular Cells","HSPC","Ery","Mk","Ery Early","Ery Late","Mk Early","Mk Late","Neutrophils 1","Neutrophils 2","Monocytes","Macrophages","Dendritic Cells (cDc1)","B Cells","T Cells","Plasma Cells"))

filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 7] = "RP Reticular Cells"
filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 20] = "WP Reticular Cells"
filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 8] = "Ery Late"
filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 12] = "Mk Early"
filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 15] = "Mk Late"
filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 18] = "Ery Early"
filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation[filtered_corrected_merged_object_sn_8patients_harmony$RNA_snn_res.1.5 == 21] = "Ery Late"

filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation = factor(filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation, levels = c("Classic Endothelial","Sinusoidal Endothelial","RP Reticular Cells","WP Reticular Cells","HSPC","Ery Early","Ery Late","Mk Early","Mk Late","Neutrophils 1","Neutrophils 2","Monocytes","Macrophages","Dendritic Cells (cDc1)","B Cells","T Cells","Plasma Cells"))

DimPlot(filtered_corrected_merged_object_sn_8patients_harmony, group.by = "complete_annotation")
```

#Zoom on lymphocytes populations to refine their annotation
```{r}
DimPlot(filtered_corrected_merged_object_sn_8patients_harmony, group.by = "simplified_annotation")
lymphocytes = subset(filtered_corrected_merged_object_sn_8patients_harmony, subset = simplified_annotation %in% c("B Cells","T Cells"))
lymphocytes <- FindVariableFeatures(lymphocytes, nfeatures = 2000)
lymphocytes <- ScaleData(lymphocytes, features = rownames(lymphocytes))
lymphocytes <- RunPCA(lymphocytes)
ElbowPlot(lymphocytes)

### Harmony & Visualisation
set.seed(1)
ndim = 10
lymphocytes <- RunHarmony(lymphocytes, group.by.vars = "orig.ident" )
lymphocytes <- RunUMAP(lymphocytes, reduction = "harmony", dims = 1:ndim, seed.use = 1)
DimPlot(lymphocytes,reduction = "umap")
lymphocytes <- RunTSNE(lymphocytes, reduction = "harmony", dims = 1:ndim, seed.use = 1)
lymphocytes <- FindNeighbors(lymphocytes, reduction = "harmony", dims = 1:ndim) %>% FindClusters( resolution = 0.5,random.seed = 1)

Markers = FindAllMarkers(lymphocytes, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
Markers %>%
    dplyr::group_by(cluster) %>%
    dplyr::slice_max(n = 5, order_by = avg_log2FC) -> top5.res

lymphocytes$lympho_annotation = lymphocytes$seurat_clusters
lymphocytes$lympho_annotation = factor(lymphocytes$lympho_annotation, levels = c(levels(tcell$lympho_annotation), "FO B Cell","MZ B Cell","CD4+ T Cell", "CD8+ T Cell", "NK Cell","Dendritic Cells (pDc)"))

lymphocytes$lympho_annotation[lymphocytes$seurat_clusters == 0] = "MZ B Cell"
lymphocytes$lympho_annotation[lymphocytes$seurat_clusters == 2] = "FO B Cell"
lymphocytes$lympho_annotation[lymphocytes$seurat_clusters == 4]= "FO B Cell"
lymphocytes$lympho_annotation[lymphocytes$seurat_clusters == 6] = "FO B Cell"
lymphocytes$lympho_annotation[lymphocytes$seurat_clusters == 1] = "CD4+ T Cell"
lymphocytes$lympho_annotation[lymphocytes$seurat_clusters == 5] = "CD8+ T Cell"
lymphocytes$lympho_annotation[lymphocytes$seurat_clusters == 3] = "NK Cell"
lymphocytes$lympho_annotation[lymphocytes$seurat_clusters == 7] = "Dendritic Cells (pDc)"

DimPlot(lymphocytes,reduction = "umap", group.by = "lympho_annotation")
```

Transferring detailed lymphocytes annotation to the complete seurat object containing all cells
```{r}
annotations <- lymphocytes$lympho_annotation
names(annotations) <- colnames(lymphocytes)

filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation = factor(filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation, levels = c(levels(filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation),"FO B Cell","MZ B Cell","CD4+ T Cell","CD8+ T Cell","NK Cell","Dendritic Cells (pDc)"))

# Map the annotations to the corresponding nuclei in the larger object
common_nuclei <- intersect(colnames(filtered_corrected_merged_object_sn_8patients_harmony), names(annotations))
filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation[common_nuclei] <- annotations[common_nuclei]

table(is.na(filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation))

DimPlot(filtered_corrected_merged_object_sn_8patients_harmony, group.by = "complete_annotation")
DimPlot_scCustom(filtered_corrected_merged_object_sn_8patients_harmony, group.by = "complete_annotation", label = TRUE, ggplot_default_colors = TRUE, label.box = FALSE, repel = FALSE, label.size  = 3.5)
```

#Zoom on HSPCs subpopulations
```{r}
HSPC_hemato = subset(filtered_corrected_merged_object_sn_8patients_harmony, subset = complete_annotation %in% c("HSPC","Ery Late","Ery Early","Mk Late","Mk Early","Neutrophils 1"))

HSPC_hemato <- FindVariableFeatures(HSPC_hemato, nfeatures = 2000)
HSPC_hemato <- ScaleData(HSPC_hemato, features = rownames(HSPC_hemato))
HSPC_hemato <- RunPCA(HSPC_hemato)
ElbowPlot(HSPC_hemato)

set.seed(1)
ndim = 10
HSPC_hemato <- RunHarmony(HSPC_hemato, group.by.vars = "orig.ident" )
HSPC_hemato <- RunUMAP(HSPC_hemato, reduction = "harmony", dims = 1:ndim, seed.use = 1)
HSPC_hemato <- FindNeighbors(HSPC_hemato, reduction = "harmony", dims = 1:ndim) %>% FindClusters( resolution = 0.7,random.seed = 1)

DimPlot(HSPC_hemato,reduction = "umap", group.by="simplified_annotation")
FeaturePlot(HSPC_hemato, features = c("CD34","MYB","MYC","GATA1","PPBP","HBB","AZU1","TRAC"), ncol=4, order = TRUE)
DimPlot(HSPC_hemato,reduction = "umap", group.by = "seurat_clusters")
Markers = FindAllMarkers(HSPC_hemato, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

Markers %>% dplyr::group_by(cluster) %>%dplyr::slice_max(n = 5, order_by = avg_log2FC) -> top5.res

HSPC_hemato$HSPC_annotation = HSPC_hemato$seurat_clusters
HSPC_hemato$HSPC_annotation = factor(HSPC_hemato$HSPC_annotation, levels = c(levels(HSPC_hemato$HSPC_annotation), "Neutrophils 1","HSPC","Proliferating Lymphoid Cells","Ery Early","Ery Late","Mk Early","Mk Late"))
HSPC_hemato$HSPC_annotation[HSPC_hemato$seurat_clusters == 0] = "Neutrophils 1"
HSPC_hemato$HSPC_annotation[HSPC_hemato$seurat_clusters == 2] = "Ery Late"
HSPC_hemato$HSPC_annotation[HSPC_hemato$seurat_clusters == 4]= "HSPC"
HSPC_hemato$HSPC_annotation[HSPC_hemato$seurat_clusters == 6] = "Mk Early"
HSPC_hemato$HSPC_annotation[HSPC_hemato$seurat_clusters == 1] = "Ery Late"
HSPC_hemato$HSPC_annotation[HSPC_hemato$seurat_clusters == 5] = "Ery Early"
HSPC_hemato$HSPC_annotation[HSPC_hemato$seurat_clusters == 3] = "Mk Late"
HSPC_hemato$HSPC_annotation[HSPC_hemato$seurat_clusters == 7] = "Ery Late"
HSPC_hemato$HSPC_annotation[HSPC_hemato$seurat_clusters == 8] = "Mk Early"
HSPC_hemato$HSPC_annotation[HSPC_hemato$seurat_clusters == 9] = "Ery Late"
HSPC_hemato$HSPC_annotation[HSPC_hemato$seurat_clusters == 10] = "Proliferating Lymphoid Cells"
DimPlot(HSPC_hemato,reduction = "umap", group.by = "HSPC_annotation")

HSPC_hemato$HSPC_annotation_simplified = HSPC_hemato$HSPC_annotation
HSPC_hemato$HSPC_annotation_simplified =  factor(HSPC_hemato$HSPC_annotation_simplified, levels = c("Neutrophils 1","HSPC","Proliferating Lymphoid Cells","Ery","Mk","Ery Early","Ery Late","Mk Early","Mk Late"))
HSPC_hemato$HSPC_annotation_simplified[HSPC_hemato$HSPC_annotation == "Mk Early"] = "Mk"
HSPC_hemato$HSPC_annotation_simplified[HSPC_hemato$HSPC_annotation == "Mk Late"] = "Mk"
HSPC_hemato$HSPC_annotation_simplified[HSPC_hemato$HSPC_annotation == "Ery Early"] = "Ery"
HSPC_hemato$HSPC_annotation_simplified[HSPC_hemato$HSPC_annotation == "Ery Late"] = "Ery"
HSPC_hemato$HSPC_annotation_simplified =  factor(HSPC_hemato$HSPC_annotation_simplified, levels = c("Neutrophils 1","HSPC","Proliferating Lymphoid Cells","Ery","Mk"))
DimPlot(HSPC_hemato,reduction = "umap", group.by = "HSPC_annotation_simplified")
```

Transferring detailed HSPC annotation to the complete seurat object containing all cells
```{r}
### Complete annotation
annotations <- HSPC_hemato$HSPC_annotation
names(annotations) <- colnames(HSPC_hemato)

filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation = factor(filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation, levels = c(levels(filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation),"Proliferating Lymphoid Cells"))

# Map the annotations to the corresponding nuclei in the larger object
common_nuclei <- intersect(colnames(filtered_corrected_merged_object_sn_8patients_harmony), names(annotations))
filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation[common_nuclei] <- annotations[common_nuclei]
table(is.na(filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation))

### Simplified annotation
annotations <- HSPC_hemato$HSPC_annotation_simplified
names(annotations) <- colnames(HSPC_hemato)

filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation = factor(filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation, levels = c(levels(filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation),"Proliferating Lymphoid Cells"))

common_nuclei <- intersect(colnames(filtered_corrected_merged_object_sn_8patients_harmony), names(annotations))
filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation[common_nuclei] <- annotations[common_nuclei]
table(is.na(filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation))
```


### IV - Exporting the clean annotated object
```{r}
filtered_corrected_merged_object_sn_8patients_harmony$final_annotation = filtered_corrected_merged_object_sn_8patients_harmony$simplified_annotation
filtered_corrected_merged_object_sn_8patients_harmony$final_annotation_high_res = filtered_corrected_merged_object_sn_8patients_harmony$complete_annotation
filtered_corrected_merged_object_sn_8patients_harmony$ident_paper = filtered_corrected_merged_object_sn_8patients_harmony$orig.ident

# Define the columns you want to keep
columns_to_keep <- c( "nCount_RNA", "nFeature_RNA", "percent.mito", "ident_paper",
  "RNA_snn_res.1.5","Disease",  "final_annotation",  "final_annotation_high_res")

DimPlot(filtered_corrected_merged_object_sn_8patients_harmony,split.by="ident_paper")
DimPlot(filtered_corrected_merged_object_sn_8patients_harmony,group.by="final_annotation_high_res")
DimPlot(filtered_corrected_merged_object_sn_8patients_harmony,group.by="final_annotation")

# Create a new object with only selected metadata columns
export <- filtered_corrected_merged_object_sn_8patients_harmony
export@meta.data <- filtered_corrected_merged_object_sn_8patients_harmony@meta.data[, columns_to_keep, drop = FALSE]

qsave(export,"updated_sn_RNAseq_human_export.qs")
```

